EDA

### 2024-11-28 other/EDA.qmd
added another graph

```{r}
df$year <- format(df$award_date, "%Y")

stacked_df <- df %>%
  filter(buyer %in% selected_buyers) %>%
  group_by(year, buyer) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

# Ensure the order of the stacks matches the specified order
stacked_df$buyer <- factor(stacked_df$buyer, levels = selected_buyers)

# Create the stacked bar chart
ggplot(data = stacked_df, aes(x = year, y = total_spending, fill = buyer)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c("National Defence" = "steelblue",
               "Public Services and Procurement Canada" = "orange",
               "Shared Services Canada" = "green",
               "Public Health Agency of Canada" = "purple",
               "Fisheries and Oceans Canada" = "red")
  ) +
  labs(
    title = "Spending by Selected Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

```
```{r}
df$year <- format(df$award_date, "%Y")

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending, fill = buyer)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  scale_fill_manual(
    values = c("National Defence" = "steelblue",
               "Public Services and Procurement Canada" = "orange",
               "Shared Services Canada" = "green",
               "Public Health Agency of Canada" = "purple",
               "Fisheries and Oceans Canada" = "red")
  ) +
  labs(
    title = "Spending by Selected Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

@@

>>>>

```{r}
# Add "year" column to the dataframe
df$year <- format(df$award_date, "%Y")

# Define selected buyers
selected_buyers <- c(
  "National Defence",
  "Public Services and Procurement Canada",
  "Shared Services Canada",
  "Public Health Agency of Canada",
  "Fisheries and Oceans Canada"
)

# Create a new column to categorize buyers
df$buyer_category <- ifelse(df$buyer %in% selected_buyers, df$buyer, "Other")

# Summarize data
stacked_df <- df %>%
  group_by(year, buyer_category) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

# Ensure the order of the stacks matches the specified order
stacked_df$buyer_category <- factor(stacked_df$buyer_category, levels = c(selected_buyers, "Other"))

# Create the stacked bar chart
ggplot(data = stacked_df, aes(x = year, y = total_spending, fill = buyer_category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c(
      "National Defence" = "steelblue",
      "Public Services and Procurement Canada" = "orange",
      "Shared Services Canada" = "green",
      "Public Health Agency of Canada" = "purple",
      "Fisheries and Oceans Canada" = "red",
      "Other" = "gray"
    )
  ) +
  labs(
    title = "Spending by Selected Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
df$year <- format(df$award_date, "%Y")

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending, fill = buyer)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  scale_fill_manual(
    values = c("National Defence" = "steelblue",
               "Public Services and Procurement Canada" = "orange",
               "Shared Services Canada" = "green",
               "Public Health Agency of Canada" = "purple",
               "Fisheries and Oceans Canada" = "red")
  ) +
  labs(
    title = "Spending by Selected Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r}
hist(df$amount,
     breaks = 30,  # Number of bins; adjust as needed
     main = "Distribution of Award Amounts",
     xlab = "Award Amount (CAD)",
     ylab = "Frequency",
     col = "skyblue",
     border = "black")
